Timeliness of security updates for Android

### 2015-03-20 securityupdates.tex
Note a paragraph we could chop if we need to
Move some figures around for security updates weekly
Fix related work to take account of the shortening of the lifetime of vulnerabilities section

The discovery dates of our vulnerabilities are not uniform. 
In particular the data shows a large gap from 2011-10-06 to 2013-02-18 where we do not know of any discoveries of critical vulnerabilities affecting all Android devices.
The cause of this quiet period is unclear.
Possible explanations are that: (i) most devices were exposed to known vulnerabilities so there was reason to look for new ones;
(ii) device manufacturers made it easier to install custom versions of Android, reducing the need for users to root their devices; and
(iii) device manufacturer specific vulnerabilities (which are not in our analysis here) were easier to find and so attackers looking for vulnerabilities adjusted their focus.

@@

\begin{figure*}
 \centering
 \includegraphics[width=\textwidth]{figures/da_norm_os}
 \caption{Android versions in \da\ data over time. The anomaly beginning in August 2014 is explained in \S\ref{sec:da_changes}.}
 \label{fig:norm_os}
\end{figure*}

@@

\begin{figure*}
 \includegraphics[width=\textwidth]{figures/nvulnerabilities_heat.pdf}
 \caption{Proportion of devices affected by different vulnerabilities. The prefix `(maybe)' indicates `maybe secure' and `(definite)' indicates `definitely insecure'. The first few vulnerabilities are all `maybe secure' as we do not have data from before that vulnerability was discovered while later ones are mostly `definitely insecure' as we know no fixing update reached the devices.
 %The change in behaviour after August 2014 is explained in \S\ref{sec:da_changes}
 }
 \label{fig:nvulnerabilities_heat}
\end{figure*}

@@

Currently, finding and exploiting vulnerabilities is typically a time-consuming manual process.
Therefore an attacker has to invest significant human resource to first find a bug and then write code to exploit the vulnerability.
To reduce the effort involved in finding a bug, attackers can look at security-related commits made to open source repositories.
Attackers have taken this approach for the Firefox web browser~\cite{Barth2011}.
Google does not release updates to the Android source code until after the release of the relevant update.
This reduces the time available for an attacker to exploit a vulnerability by looking at security updates disclosed in the source code repository from the period (\textbf{first fix}, \textbf{fix deployed}) to (\textbf{fixing release}, \textbf{fix deployed}), unfortunately this can still take months .
For example, the zergRush vulnerability was fixed in a release before it was publicly discovered but it still took 27 months for 90\% of the devices in our study to be definitely fixed\footnote{Definitely fixed is 1 $-$ (definitely insecure + maybe secure), maybe fixed is 1 $-$ definitely insecure} (25 months for 90\% of devices maybe fixed).

Brumley et al. has shown it is possible to automatically generate exploits from binary fixes, for example by using security updates issued by manufacturers~\cite{Brumley2008}. 
Similarly, the Woodpecker tool automatically finds permission leaks in stock Android phone images~\cite{Grace2012}.
Once such automated techniques mature, it will become even more important to provide security updates promptly to every Android device in the ecosystem because an attacker can turn the contents of a \textbf{fixing release} into an exploit in a matter of minutes or hours.

>>>>

The discovery dates of our vulnerabilities are not uniform. 
In particular the data shows a large gap from 2011-10-06 to 2013-02-18 where we do not know of any discoveries of critical vulnerabilities affecting all Android devices.
The cause of this quiet period is unclear.
Possible explanations are that: (i) most devices were exposed to known vulnerabilities so there was reason to look for new ones;
(ii) device manufacturers made it easier to install custom versions of Android, reducing the need for users to root their devices; and
(iii) device manufacturer specific vulnerabilities (which are not in our analysis here) were easier to find and so attackers looking for vulnerabilities adjusted their focus.
\todolater{If we need to save space then delete this paragraph}

@@

\begin{figure}[h]
 \centering
 \includegraphics[width=\textwidth]{figures/da_norm_os}
 \caption{Android versions in \da\ data over time. The anomaly beginning in August 2014 is explained in \S\ref{sec:da_changes}.}
 \label{fig:norm_os}
 \includegraphics[width=\textwidth]{figures/nvulnerabilities_heat.pdf}
 \caption{Proportion of devices affected by different vulnerabilities. The prefix `(maybe)' indicates `maybe secure' and `(definite)' indicates `definitely insecure'. The first few vulnerabilities are all `maybe secure' as we do not have data from before that vulnerability was discovered while later ones are mostly `definitely insecure' as we know no fixing update reached the devices.
 %The change in behaviour after August 2014 is explained in \S\ref{sec:da_changes}
 }
 \label{fig:nvulnerabilities_heat}
\end{figure}

@@

@@

Currently, finding and exploiting vulnerabilities is typically a time-consuming manual process.
Therefore an attacker has to invest significant human resource to first find a bug and then write code to exploit the vulnerability.
To reduce the effort involved in finding a bug, attackers can look at security-related commits made to open source repositories.
Attackers have taken this approach for the Firefox web browser~\cite{Barth2011}.
Google does not release updates to the Android source code until after the release of the relevant update.
This reduces the time available for an attacker to exploit a vulnerability by looking at security updates disclosed in the source code repository.
Reducing it from, the interval between the commit of the first fix and the fix being deployed to most devices, to the interval between the first fixing release and the fix being deployed to most devices; unfortunately this can still take months or years.
For example, the zergRush vulnerability was fixed in a release before it was publicly discovered but it still took 27 months for 90\% of the devices in our study to be definitely fixed\footnote{Definitely fixed is 1 $-$ (definitely insecure + maybe secure), maybe fixed is 1 $-$ definitely insecure} (25 months for 90\% of devices maybe fixed).

Brumley et al. has shown it is possible to automatically generate exploits from binary fixes, for example by using security updates issued by manufacturers~\cite{Brumley2008}. 
Similarly, the Woodpecker tool automatically finds permission leaks in stock Android phone images~\cite{Grace2012}.
Once such automated techniques mature, it will become even more important to provide security updates promptly to every Android device in the ecosystem because an attacker can turn the contents of a released fix into an exploit in a matter of minutes or hours.
